Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-06 trunk/paper.tex

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$.

Our primary reconfigurable design exploits set associativity in conventional
cache organizations. Figure~\ref{fig:reconfig} depicts the block diagram for the
organization of our design. An n-way set associative cache consists of $n$ data
and tag arrays. We divide the each level of on-chip memory into partitions at
the granularity of the $k$-ways, where $k$ is determined by the number of
available capacity values for a specific memory technology on the provided
bandwidth curve. For example, if there are 8 available capacities for STT-RAM
(from 2MB to 1GB), we will determine the granularity by $k=n/8$.The desirability
of such configuration is to exploit the division of ways already present in a
conventional cache organization.

The index part of the input address is used to index all the
ways of the data and tag array. The tag part of the input address is sent to the
comparators of all the ways to determine if there is a match with any of the
tags read from the tag array. If there is a match, a hit is signaled on the
valid output line and data from the corresponding way of the data array is sent
onto the output data lines.

reconfiguration support: how reconfigurable (Li -> Li+1 (Li+0.5); hardware
modification to support reconfiguration.

reconfigure the unused capacity:

- if (> si): becomes si+1

- else: unused (or becomes lut)

@@

statistical prediction

>>>>

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$. Furthermore, the rest of the capacity can be
reconfigured to become level-i.5 if necessary. We describe the reconfiguration
algorithm as follows.

<pseudocode>

Our primary reconfigurable design exploits set associativity in conventional
cache organizations. The desirability of such configuration is to exploit the
division of ways already present in a conventional cache organization.
Figure~\ref{fig:reconfig} depicts the block diagram for the organization of our
design. An n-way set associative cache consists of $n$ data and tag arrays. We
divide the each level of on-chip memory into partitions at the granularity of
the $k$-ways, where $k$ is determined by the available capacity range of a
specific memory technology on the provided bandwidth curve. <100+ ways?
experiment section> Reconfiguration will not affect the bits of the address
fields that are used as tag, index, and block offset bits. Modifications to the
conventional cache architecture include:

\begin{itemize}
\item \textbf{Memory status vector} At each memory level, the corrent
configuration is stored in a memory status vector.
\item \textbf{Input and output paths} The input and output data paths are
duplicated. 
\item \textbf{Additional multiplexors} Futhermore, additional wiring and
multiplexors at address decoders and tag comparators are also required to
support the reconfiguration.
\end{itemize}

@@

Memory reconfiguration relies on accurate predictions of demand bandwidth of a
workload in response to the dynamically varying application characteristics.
Conventional last value or table based predictors can model neither long range
patterns of an application nor patterns with variable lengths. In this work, we
employ a statistical predictor to support for the BAROM design. The basic idea
of the predictor is borrowed from language modeling~\cite{smoothing-tech}.

hardware: global predictor sitting before caches
statistical prediction
